\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{tgpagella}

\input{preamble.tex}

\title{MATH 342}
\date{Winter Term 1 2024}
\begin{document}
\maketitle

\begin{tcolorbox}[title=,
    fonttitle=\huge\sffamily\bfseries\selectfont,interior style={left
    color=contcol1!40!white,right color=contcol2!40!white},frame
    style={left color=contcol1!80!white,right
    color=contcol2!80!white},coltitle=black,top=2mm,bottom=2mm,left=2mm,right=2mm,drop
  fuzzy shadow,enhanced,breakable]
  \tableofcontents
\end{tcolorbox}

\newpage

\section{Linear Codes}
\begin{definition}
  A \textbf{linear code} \( C \) is a subspace of \( V(n, q) \) for
  some positive \( n \). Thus, \( C \) is linear iff
  \begin{enumerate}
    \item \( u, v \in C \implies u + v \in C \)
    \item \( u \in C, a \in GF(q) \implies au \in C \)
  \end{enumerate}

  If \( C \) is a \( k \)-dimensional subspace of \( V(n, q) \) then
  \( C \) is called an \( [n, k] \) or \( [n, k, d] \) code.
\end{definition}
\begin{remark}
  A \( q \)-ary \( [n, k, d] \) code is also a \( q \)-ary \( (n,
  q^k, d) \) code but converse is not true.

  \( 0 \) must be in \( C \).
\end{remark}
\begin{definition}
  The \textbf{weight} of a vector \( x \) is defined to be the number
  of non-zero entries of \( x \).
\end{definition}
\begin{lemma}
  If \( x, y \in V(n, q) \) then \(
  d(x, y) = w(x - y) \).
\end{lemma}
\begin{proof}
  The vector \( x-y \) has non-zero entries in those places where \(
  x, y \) differ.
\end{proof}
\begin{theorem}
  Let \( C \) be linear. Then \[
    d(C) = \min_{x \in C}w(x)
  \]
\end{theorem}
\begin{proof}
  There are codewords \( x, y \) such that \( d(x, y) = c = d(C) \).
  (As otherwise the distance of the codeword could never be \( c \)).

  Then, \( d(C) = w(x - y) \ge \min_{x \in C}w(x) \) since \( x-y  \)
  is a codeword of \( C \). However, for some \( x \in C \)  \(
  \min_{x \in C}w(x) = w(x) = d(x, 0) \ge d(C)\). Thus have both inequalities.
\end{proof}
\begin{definition}
  A \( k \times n \) matrix whose rows form a basis of a linear \(
  [n, k] \) code is called a \textbf{generator matrix} of the code.
\end{definition}
\begin{theorem}
  Let \( G \) be a generator matrix of an \( [n, k] \) code. Using
  EROs, \( G \) can be transformed into \textbf{standard form} \[
    [I_k \mid A]
  \] where \( I_k \) is the \( k \times k  \) identity and \( A is k
  \times (n-k) \)
\end{theorem}
\subsection{Encoding with a Linear Code}
\begin{definition}
  Let \( C \) be \( [n, k] \)-code over \( GF(q) \) with generator \(
  G \). \( C \) contains \( q^k \) codewords, so that is the max
  possible number of distinct messages.

  A message is a \( k \)-tuple of \( V(k, q) \). Encode a message
  vector \( u = u_1 u_2 \dots u_k  \) by multiplying as \( uG \).

  Note that this is a map \( V(k, q) \to C \subset V(n, q) \)
\end{definition}
\begin{corollary}
  If \( G \) is in standard form, then the encoding \( x = uG = (x_1,
  x_2, x_3 ,\dots, x_k, x_{k+1}, \dots, x_n)  \) has \( x_i = u_i \)
  for \( 1 \le i \le k \) (called \textbf{message digits}) and \(
  x_{k+i} = \sum_{j=1}^k a_{ji}u_j\) for \( 1 \le i \le  n-k \)
  (called \textbf{check digits}).
\end{corollary}
\begin{example}
  Let \( C \) be binary \( [7, 4] \) code. Let \[
    G =
    \begin{bmatrix}
      1 & 0 & 0 & 0 & 1 & 0 & 1 \\
      0 & 1 & 0 & 0 & 1 & 1 & 1 \\
      0 & 0 & 1 & 0 & 1 & 1 & 0 \\
      0 & 0 & 0 & 1 & 0 & 1 & 1
    \end{bmatrix}
  \] be its generator matrix.

  A message vector \( u_1, u_2, u_3, u_4 \) is encoded as \[
    (u_1, u_2, u_3, u_4, u_1 + u_2 + u_3, u_2 + u_3 + u_4, u_1 + u_2 + u_4)
  \]

  For instance \( 1000 \) is encoded as \( 1000101 \).
\end{example}
\subsection{Decoding with a Linear Code}
\begin{definition}
  Suppose the codeword \( x \) is sent and the codeword \( y \) is
  recieved. Define the \textbf{error vector} \( e \) as \( e = y-x \).
\end{definition}
\section{Cyclic Codes}
\begin{definition}
  A code \( C \) is cyclic if
  \begin{enumerate}
    \item \( C \) is linear.
    \item \( C \) is closed under shift \( S \), i.e., \( w \in C
      \implies S(w) \in C \).
  \end{enumerate}
\end{definition}
\begin{example}
  The code \( C = \{000, 101, 011, 110\} \) is cyclic.
\end{example}
\begin{remark}
  Note that a shift is a right shift, but left shifts can be
  simulated by shifting right \( n-1 \) times where \( n \) is the
  length of the codeword.
\end{remark}
\begin{remark}
  Can view a cyclic code as a polynominal, where the digits of the
  codeword are coefficients for a polynominal of degree \( n-1 \), \(
  a_0 + a_1x + a_2x^2 + \cdots + a_{n-1} x^{n-1}  \).
\end{remark}
\begin{definition}
  In a cyclic code, declare \( x^n \equiv 1 \iff x^n-1 = 0 \).

  Then a cyclic code \( C \) is a subspace of  \( Z_p[x]/(x^n-1) \)
  such that \( C \) is closed under multiplication with \( x \).

\end{definition}
\subsection{The Ring of Polynomials}
\begin{definition}
  Let \( F \) be a field.
  Then \( F[x] \) is the set of polynominals with coefficients in \(
  F \). Let \( \deg f = d \), and \( f \) is \textbf{monic} if the
  term with highest degree has coefficient one.
\end{definition}
\begin{definition}
  Define division as \[
    \frac{f(x)}{g(x)} = q(x) + \frac{r(x)}{g(x)}
  \] where \( \deg r < \deg g \).
\end{definition}
\begin{definition}
  \( g \) divides \( f \) if \( \frac{f(x)}{g(x)}  \) has \( r(x) = 0 \).
\end{definition}
\begin{definition}
  Let \( f(x)\) be a fixed polynominal in \( F[x] \). Two polynominals \( g, h \) are said to be \textbf{congruent} modulo \( f \) symbolized by \( g(x) = h(x) \pmod{f(x)} \) if \( g(x) - h(x) \) is divisible by \( f(x) \).
\end{definition}
\begin{remark}
  Any polynominal \( a(x) \) is congruent modulo \( f(x) \) to a unique polynominal \( r(x) \), which is the principal remainder when \( a(x) \) is divided by \( f(x) \).
\end{remark}
\begin{definition}
  The GCD of \( f, g \) is the monic polynominal of highest degree
  that divides them.
\end{definition}
\begin{definition}
  \( \alpha \in  F   \) is a \textbf{root} of \( f \) if \( f(\alpha   ) = 0\)
\end{definition}
\begin{theorem}
  \( \alpha  \) is a root of \( f \) if and only if \( x - \alpha  \)
  divides \( f \).
\end{theorem}
\begin{corollary}
  If \( \deg f = n \), then \( f \) can have at most \( n \) roots.
\end{corollary}
\begin{definition}
  \( f(x) \in F[x] \) is \textbf{irreducible} if \( f(x) \neq
  g(x)h(x) \) where \( \deg g, \deg h < \deg f \). Usually take
  irreducibles monic.
\end{definition}
\begin{theorem}
  Every \( f \) can be factored as the product of irreducibles.
\end{theorem}
\begin{corollary}
  Irreducible degree 3 or less polynominals must have no roots.
\end{corollary}
\begin{remark}
  To find all monic irreducibles, helpful to list all polynominals
  (There are \( p^{n} \) of them, where \( p \) is the modulus, and
  \( n \) the degree), then count the reducible ones and subtract.
  Use stars and bars formula.
\end{remark}
\begin{definition}
  \( Z_p[x]/ f(x) = \{\text{all principal remainders divided by }f\}
  =  \{\text{all }r(x) \text{ such that } \deg r < \deg f \} \) which
  is the set \( \{a_0 + a_1x + \cdots + a_{n-1}x^{n-1} \}  \).
\end{definition}
\begin{example}
  The ring \( Z_2[x]/(x^3+x+1) \) has the elements \( \{0, 1, x, x+1,
  x^2, x^2+1, x^2 + x, x^2 + x + 1 \}\).
\end{example}
\subsubsection{Inverses}
\begin{definition}
  \( g^{-1} \) is the element such that \( gg^{-1} = 1 \) in \(
  Z_p[x]/f(x) \). Note that the inverse may not always exist.
\end{definition}
\begin{theorem}
  \( g^{-1} \) exists if and only \( \gcd(f, g) = 1 \) in a ring mod \( f \).
\end{theorem}
\begin{theorem}
  \( Z_p[x]/f(x) \) is a field if and only if \( f(x) \) is
  irreducible in \( Z_p[x] \)
\end{theorem}
\begin{example}
  \( Z_2[x] /(x^2+x+1) \) is a field with four elements, also called
  \( \mathbb{F}_4 \).

  \( Z_2[x]/(x^3+x+1) \) is a field with eight elements, also called
  \( \mathbb{F}_8 \).
\end{example}
\begin{proposition}
  Every field with finite number of elements has form \( \mathbb{F}_q
  \) where \( q = p^n \) for some prime \( p \).
\end{proposition}
\begin{remark}
  If two fields have the same modulus \( p \) and the same degree
  modular polynominal, then they are isomorphic.
\end{remark}
\begin{proposition}
  Every \( \mathbb{F}_{p^n} \) has a primitive element \( g \) such
  that the powers \( g, g^2, \dots, g^{p^n-1} = 1  \) are distinct.
\end{proposition}
\subsection{Ideals}
\begin{definition}
  Let \( R_n = F[x]/(x^n-1) \) where \( F = F_q \) be implicit.
\end{definition}
\begin{definition}\label{baz}
  A simpler definition for a cyclic code \( C \subset R_n\) is
  \begin{enumerate}
    \item \( 0 \in C \)
    \item \( g(x), h(x) \in  C \implies g(x)+h(x) \in  C \)
    \item \( g(x) \in C, r(x) \in \R_n \implies  r(x)g(x) \in C \)
  \end{enumerate}
  \( C \) is called an \textbf{ideal} in the ring \( R_n \).
\end{definition}
\begin{proof}
  Suppose \( C \) is cyclic in \( R_n \). \( C \) is thus linear and so the additive closure condition holds.

  Let \( a(x) \in C \) and \( r(x) = r_0 + r_1x + \cdots + r_{n-1}x^{n-1} \in R_n   \). Since multiplication by \( x  \) corresponds to a cyclic shift, we have \( xa(x) \in  C \) and \( x^2a(x) \in C \) and so on. Hence \( r(x)a(x) = r_0a(x) + r_1xa(x) + \cdots + r_{n-1}x^{n-1}a(x)  \) is also in \( C \) since each term is in \( C \). This the multiplicative closure condition holds.

  Taking \( r(x) = 0 \) satifies the zero condition.
\end{proof}
\begin{remark}
  Note that taking \( r(x) = c \) in the above proof implies \( C  \) is linear and \( r(x) = x \) implies \( C \) is cyclic.
\end{remark}
\begin{definition}
  Let \( R  \) be a \href{https://en.wikipedia.org/wiki/Ring_(mathematics)}{ring}.
  \( I \) is \textbf{principal} if \( I = \langle g \rangle = R \cdot g \).
\end{definition}
\begin{example}
  \( R = \Z, I = \langle 12, 18 \rangle \), \( I = \{a \cdot 12 + b
  \cdot  18 \mid a, b \in  \Z \} \).
  Claim \( 6 \in I \), since \( 6 \) is the gcd.
  Claim \( I = \langle 6 \rangle  \) because \( 12a + 18b = \Z \cdot 6 \)
\end{example}
\begin{proposition}
  Every ideal in \( \Z \) is principal. \( \langle g_1, g_2, \dots,
  g_n \rangle = \langle d \rangle  \) where \( d = \gcd(g_i) \).
\end{proposition}
\begin{example}
  Given an ideal \( I \subset \Z \), know that \( I = \langle d
  \rangle  \). Find \( d \) by taking the smallest positive number in \( I \).
\end{example}
\begin{theorem}
  Every cyclic code (non-zero) is of the form \( \langle g(x) \rangle
  \subset R_n \) where \( g \) divides \( x^n-1 \). The generator \(
  g \) is unique if it is monic.
\end{theorem}
\begin{corollary}
  Cyclic codes are 1-1 with monic \( g \) that divide \( x^n-1  \)
\end{corollary}

\subsection{Generator Polynomials}
\begin{definition}
  \[
    \langle f(x) \rangle = \{r(x)f(x) \mid r(x) \in R_n\}
  \]
\end{definition}
\begin{theorem}
  For any \( f(x) \in R_n \), the set \( \langle f(x) \rangle \) is a cyclic code,
  and it is called the code generated by \( f(x) \).
\end{theorem}
\begin{proof}
  Check the conditions of \autoref{baz}
  \begin{enumerate}
    \item Let \( a(x)f(x), b(x)f(x) \in \langle f(x) \rangle \). then \( a(x)f(x) + b(x)f(x) = (a(x)+b(x))f(x) \in \langle f(x) \rangle  \)
    \item Let \( a(x)f(x) \in \langle f(x) \rangle\), \( r(x) \in R_n \), then \[
        r(x)(a(x)f(x)) = (r(x)a(x)) \in \langle f(x) \rangle
      \]
  \end{enumerate}
\end{proof}
\begin{example}
  Consider the code \( C = \langle 1 + x^2 \rangle \) in \( R_3 \)
  with \( F = GF(2) =\Z_2\). Multiplying by each of the 8 elements of \(
  R_3 \), (i.e. \( 0, 1, x, x+1, x^2, x^2+1, x^2 + x, x^2+x+1 \)) and reducing modulo \( x^3-1 \) produces only 4 distinct
  codewords, namely \( 0, 1+x, 1+x^2, x+x^2 \). Thus \( C \) is the
  code \( \{000, 110, 101, 011\} \).
\end{example}
\begin{theorem}\label{bar}
  Let \( C  \) be a non-zero cyclic code in \( R_n \). Then,
  \begin{enumerate}
    \item There exists a unique monic polynominal \( g(x)\) of
      smallest degree in \( C \).
    \item \( C = \langle g(x) \rangle  \)
    \item \( g(x) \) is a factor of \( x^n-1 \)
  \end{enumerate}
\end{theorem}
\begin{definition}
  The polynominal given by the above is called the \textbf{generator
  polynominal} of \( C \).
\end{definition}
\begin{example}
  We find all the binary cyclic codes of length \( n=3 \). Factor \( x^3-1 \) as \( (x-1)(x^2+x+1) \), both irreducible. By \autoref{bar}. Thus, a list of binary cyclic codes is:
  \begin{enumerate}
    \item Generator Polynomial: \( 1 \), Code in  \( R_3 \): All of \( R_3 \), Corresponding Code in \( V(3, 2) \): all of \( V(3, 2) \)
    \item Generator Polynomial: \( x+1 \), Code in \( R_3 \): \( \{0, 1+x, x+x^2, 1+x^2 \}\), Corresponding Code in \( V(3, 2) \): \( \{000, 110, 011, 101\} \).
    \item Generator Polynomial: \( x^2+x+1 \), Code in \( R_3 \): \( \{0, 1+x+x^2\}\), Corresponding Code in \( V(3, 2) \): \( \{000, 111\} \).
    \item Generator Polynomial: \( x^3-1 (=0) \), Code in \( R_3 \): \( \{0\} \), Corresponding Code in \( V(3, 2) \): \( \{000\} \).
  \end{enumerate}
\end{example}
\begin{lemma}
  Let \( g(x) = g_0 + g_1x + \cdots + d_rx^r \) be generator
  polynominal of a cyclic code. Then \( g_0 \neq 0 \).
\end{lemma}
\begin{example}
  Let \( p=2 \), \( n=7 \) (binary codes of length 7). Find all cyclic codes. Must find all \( g(x) \mid x^n-1 \) since they will be generators (also multiplied togeher). Have \( x^n-1 = (x-1)(x^3+x+1)(x^3+x^2+1) \). Let \( a(x), b(x), c(x) \) correspond to these three polynominals. Then,
  \begin{center}
    \[
      \begin{array}{ c|c|c }
        \deg g & g(x) & \dim C \\
        \hline
        \hline
        0 & 1 & 7 \\
        1 & a(x) & 6\\
        2 & - & 5\\
        3 & b(x), c(x) & 4\\
        4 & a(x)b(x), a(x)c(x) & 3\\
        5 & - & 2\\
        6 & b(x)c(x) & 1\\
        7 & a(x)b(x)c(x) & 0\\
      \end{array}
    \]
  \end{center}

  Note that \( \dim C = n - \deg g(x) \).

  Confused here and next theorem.
\end{example}

\subsection{Generator and Parity Check Matrices}
\begin{theorem}
  Let \( C = \langle g(x) \rangle\).
  Then every codeword in \( C \) can be written as \( w(x) = a(x)g(x) \) where \( \deg a(x) < n - \deg g(x) \). Moreover, such \( a  \) is unique.
\end{theorem}
\begin{example}
  Continued from previous example. Pick \( g(x) = (x+1)(x^3+x+1) \). Note that \( \deg g = 4 \). Then any codeword \( w(x) = (a_0 + a_1x + a_2x^2)g(x) = a_0g + a_1xg + a_2x^2g \). Thus \( g, xg, x^2g \) are a basis for \( C \).
\end{example}
\begin{theorem}
  Let \( C = \langle g \rangle \). Let \( \deg g = d \).  Then \( g, xg, x^2g \dots x^{n-d-1}g  \) form a basis for \( C  \). There are \( n-d  \) elements.
\end{theorem}
\begin{theorem}\label{foo}
  Let \( C  \) be cyclic with generator \[
    g(x) = g_0 + g_1x + \cdots  + g_rx^r
  \] of degree \( r  \). Then \( \dim C = n-r \) with generator matrix \[
    G =
    \begin{bmatrix}
      g_0 & g_1 & g_2 & \cdots&  \cdots & g_r & 0 & 0 & 0\\
      0 & g_0 & g_1 & g_2 & \cdots&  \cdots & g_r & 0 & 0\\
      0 & 0 & \vdots & \vdots & \vdots & \cdots&  \cdots & \ddots & 0\\
      0 & 0 & 0 & g_0 & g_1 & g_2 & \cdots&  \cdots & g_r\\
    \end{bmatrix}
  \]
  This matrix is \( n-r\times n \).
\end{theorem}
\begin{example}
  12.13
\end{example}
\begin{example}
  Let \( n=12 \) and work in \( GF(2) \). How many cyclic codes of dimension \( k=8 \)?

  This is the same as saying how many \( g(x) \) of degree 4 since \( \dim C = n - \deg g \).

  Factor \( x^{12} - 1 \) into \( (x+1)^4(x^2+x+1)^4 \).

  Thus there are 3 such codes comprised of the polynominals \( (x+1)^4, (x^2+x+1)^2, (x^2+x+1)(x+1)^2 \).
\end{example}
\begin{definition}
  Let \( C \) be cyclic \( [n, k] \)-code with generator \( g(x) \).
  By theorem, \( g(x) \) is a factor of \( x^n-1 \) and so \[
    x^n-1= g(x)h(x)
  \] for some polynominal \( h \). Note that \( h  \) is monic since
  \( g  \) is. \( g(x) \) has degree \( n-k  \) from \autoref{foo} so
  \( h  \) has degree \( k \). The polynominal \( h  \) is called the
  \textbf{check polynominal} of \( C \).
\end{definition}
\begin{theorem}
  Suppose \( C \) is cyclic in \( R_n  \) with generator \( g  \) and
  check polynominal \( h \). Then an element \( c(x)  \) is a
  codeword of \( C \) iff \( c(x)h(x) = 0 \).
\end{theorem}
\begin{example}
  Let \( n = 7 \) working in \( GF(2) \), with \( g(x) =  (x-1)(x^3+x+1)\). Then \( h(x) = x^3+x^2+1 \). Want to know if \( w(x) = x^6+x^3+x^2+x \in C \). Multiplying with \( h(x) \) yields \( 0 \) so it is a codeword.
\end{example}
\begin{theorem}
  Suppose \( C  \) is cyclic \( [n, k] \)-code with check polynominal
  \( h(x) = h_0 + h_1x + \cdots + h_kx^k \). Then,
  \begin{enumerate}
    \item a parity-check matrix for \( C  \) is \[
        H =
        \begin{bmatrix}
          h_k & k_{k-1} & \cdots  & h_0 & 0 & 0 & 0 & 0\\
          0 & h_k  & k_{k-1} & \cdots  & h_0 & 0 & 0 & 0\\
          0 & 0 & h_k  & k_{k-1} & \cdots  & h_0 & 0 & 0\\
          0 & 0 & 0 & h_k  & k_{k-1} & \cdots  & h_0 & 0\\
          0 & 0 & 0 & 0 & h_k  & k_{k-1} & \cdots  & h_0\\
        \end{bmatrix}
      \]
    \item \( C^\perp \) is a cyclic code generated by the polynominal \(
      \overline{h}(x) = h_k + h_{k-1}x + \cdots + d_0x^k   \)
  \end{enumerate}

  The size of \( H \) is \( n - k \times n\) since \( n-k = \dim C^\perp = \deg g  \) 
\end{theorem}
\begin{definition}
  The polynominal \( \overline{h}(x) = x^kh(x-1) = h_k + h_{k-1}x + \cdots + h_0x^k    \) is called the \textbf{reciprocal polynominal} of \( h \), its coefficients are those of \( h   \) in reverse order.

  We may regard \( \overline{h}  \) as the generator of \( C^\perp  \) though we should multiply by \( h_0^-1 \) to make it monic.
\end{definition}
\begin{remark}
  The polynominal \( h(x-1) =x^{n-k}\overline{h}(x) \) is a member of \( C^\perp  \)
\end{remark}
\end{document}
